%-----------------------------------------------------------------------------
% Author: Rayla Kurosaki
% GitHub: https://github.com/rkp1503
%-----------------------------------------------------------------------------
\documentclass{rayla-project}
\usepackage{rayla-math-style}

%-----------------------------------------------------------------------------
% Project Information
%-----------------------------------------------------------------------------
\myUniversity{
    Rochester Institute of Technology\\
    College of Science\\
    School of Mathematical Sciences
    }
\myTitle{Complex Dynamics of a Three Species Ecosystem}
\myName{Ramsey (Rayla) Phuc}
\courseid{MATH-761.01}
\courseName{Mathematical Biology}
\professorName{Dr. Ephraim Agyingi}
\term{2022/08/22 - 2022/12/05}
\dueDate{2022/12/12}

%----------------------------------------------------------------------------- 
% Start of Document
%-----------------------------------------------------------------------------
\begin{document}

    \maketitle

    %-------------------------------------------------------------------------
    % Abstract
    %-------------------------------------------------------------------------
    \begin{abstract}
        This paper investigates the dynamics of a model system by modifying key assumptions to explore alternative ecological interactions. The original model, outlined in the paper \cite{PANJA2022100153}, initially assumes a competition interaction between species $X$ and $Y$. However, we consider the implications of assuming a mutualism interaction instead. Additionally, we modify the assumption of a commensalism interaction between species $X$ and $Z$ to examine the consequences of an amensalism interaction. Furthermore, the model is expanded by assuming a logistic growth pattern for species Z. Through these modifications, we determine the new equilibrium points and conduct a comprehensive analysis of their characteristics. Numerical computations are performed to assess the stability or instability of the newly derived equilibrium points. The obtained results offer valuable insights into the dynamics and stability of the modified model, shedding light on the consequences of mutualism and amensalism interactions in the examined ecological system. This research contributes to a deeper understanding of the interrelationships between species interactions and population dynamics in ecological systems.
    \end{abstract}

    %-------------------------------------------------------------------------
    % Body
    %-------------------------------------------------------------------------
    \input{contents/introduction.tex}
    \input{contents/original-model.tex}
    \input{contents/modified-model.tex}
    \input{contents/identifying-equilibria.tex}
    \input{contents/stability-analysis.tex}
    \input{contents/numerical-simulations.tex}

    %-------------------------------------------------------------------------
    % Bibliography
    %-------------------------------------------------------------------------
    \newpage
    % \nocite{*}
    \printbibliography[title={Bibliography}]

    %-------------------------------------------------------------------------
    % Appendix
    %-------------------------------------------------------------------------
    \newpage
    \appendix
    \input{appendices/figures.tex}

\end{document}
